---
output: github_document
bibliography: bibliography.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/poissonconsulting/ypr.svg?branch=master)](https://travis-ci.org/poissonconsulting/ypr)
[![Coverage status](https://codecov.io/gh/poissonconsulting/ypr/branch/master/graph/badge.svg)](https://codecov.io/github/poissonconsulting/ypr?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)

# ypr

## Introduction

`ypr` is an R package that implements equilibrium-based yield per recruit methods [@walters_fisheries_2004].

The yield can be based on the number of fish caught (or harvested) or biomass for all fish or just large (trophy) individuals.

The key life history parameters are

- The growth coefficient (`k`) and mean maximum length (`Linf`) from the Von Bertalanffy growth curve
- The length at which 50% spawning (`Ls`)
- The length at which 50% vulnerable to harvest (`Lv`)
- The number of spawners per spawner at low density (`Rk`)

For definitions of all `r length(ypr::ypr_population())` possible parameters see the help for `ypr_population()`.
For an explanation of the calculations see the vignette `ypr`.

The calculations do not account for parameter uncertainty, environmental fluctuations, predator-prey dynamics, angler responses or density-dependent growth.

## Demonstration

```{r, fig.width=6, fig.height=4}
library(ypr)
population <- ypr_population(Rk = 10)
ypr_plot_yield(population)
ypr_tabulate_yield(population)
```

## Installation

To install the latest development version from [GitHub](https://github.com/poissonconsulting/ypr)
```
# install.packages("devtools")
devtools::install_github("poissonconsulting/ypr")
```

To install the latest development version from the Poisson drat [repository](https://github.com/poissonconsulting/drat)
```
# install.packages("drat")
drat::addRepo("poissonconsulting")
install.packages("ypr")
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/ypr/issues).

[Pull requests](https://github.com/poissonconsulting/ypr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms.

## References
