---
output: github_document
bibliography: bibliography.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/poissonconsulting/ypr.svg?branch=master)](https://travis-ci.org/poissonconsulting/ypr)
[![Coverage status](https://codecov.io/gh/poissonconsulting/ypr/branch/master/graph/badge.svg)](https://codecov.io/github/poissonconsulting/ypr?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)

# ypr

## Introduction

`ypr` is an R package that implements yield per recruit methods [@walters_fisheries_2004].

The yield, which is provided in terms of the proportion of the unfished population,
can be based on the total biomass vs number of fish,
harvested vs captured and large (trophy) versus all fish.

The yield can also be calculated with lower and upper limits that represent the uncertainty in 
one or more life history parameters.

The key life history parameters are

- The growth coefficient (`k`) and mean maximum length (`Linf`) from the Von Bertalanffy growth curve
- The condition scaling exponent (`b`) from the weight-length relationship
- The length at maturity (`Lm`)
- The length at which 50% vulnerable to harvest (`Lv`)
- The number of spawners per spawner at low density (`Rk`)

The calculations do not account for environmental fluctuations, predator-prey dynamics, angler responses or density-dependent growth.

## Demonstration

```{r}
library(ypr)
population <- ypr_population(Rk = 3)
ypr_schedule(population)
ypr_plot(population)
```

## Installation

To install the latest development version from [GitHub](https://github.com/poissonconsulting/ypr)
```
# install.packages("devtools")
devtools::install_github("poissonconsulting/ypr")
```

To install the latest development version from the Poisson drat [repository](https://github.com/poissonconsulting/drat)
```
# install.packages("drat")
drat::addRepo("poissonconsulting")
install.packages("ypr")
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/ypr/issues).

[Pull requests](https://github.com/poissonconsulting/ypr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). 
By participating in this project you agree to abide by its terms.

## References
