---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Travis build status](https://travis-ci.org/poissonconsulting/ypr.svg?branch=master)](https://travis-ci.org/poissonconsulting/ypr)
[![Coverage status](https://codecov.io/gh/poissonconsulting/ypr/branch/master/graph/badge.svg)](https://codecov.io/github/poissonconsulting/ypr?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)

# ypr

ypr is an implementation of yield per recruit methods for 
recreational fisheries in R.
Given a list of life history parameters it can generate a data frame
of the schedule and find the capture rate which maximises the yield. 
The yield, which is provided in terms of the proportion of the unfished population,
can be based on the total biomass vs number of fish; 
harvested vs captured and large (trophy) versus all fish.

## Demonstration

```{r}
library(ypr)
population <- ypr_population()
ypr_schedule(population)
ypr_plot(population)
```


## Assumptions and Limitation

The calculations assume that growth follows a von Bertalanffy growth curve;
maturation is length based;
natural mortality is constant;
vulnerability to the fishery is knife-edged;
and capture, release and hooking mortality are constant.
It also assumes that a released individual cannot be recaught in the same year.
The results suffers from the same limitations as all yield per recruit methods.

## Installation

To install the latest development version from [GitHub](https://github.com/poissonconsulting/ypr)
```
# install.packages("devtools")
devtools::install_github("poissonconsulting/ypr")
```

To install the latest development version from the Poisson drat [repository](https://github.com/poissonconsulting/drat)
```
# install.packages("drat")
drat::addRepo("poissonconsulting")
install.packages("ypr")
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/ypr/issues).

[Pull requests](https://github.com/poissonconsulting/ypr/pulls) are always welcome.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

