---
title: "Ecotypes"
author: "Joe Thorley and Ayla Pearson"
date: "`r Sys.Date()`"
bibliography: bibliography.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ecotypes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 4
)
```


```{r, echo=FALSE}
library(ypr)
```

## Ecotypes

In the ypr package a population can consist of more than one ecotype.
A population is considered to be group of panmictic fish that are indistinguishable to anglers
while ecotypes are considered to be subsets of a population that follow alternative life-history strategies.
Consequently, ecotypes in the same population must share the key stock recruitment (`BH`, `RK`, `tR` and `Rmax`) and fishery (`pi`, `Llo`, `Lup`, `Nc`, `rho`, `Hm` and `q`) parameters.
Finally, in order to use a yield-per-recruit approach it is necessary to assume that the relative proportion of recruits adopting each life-history strategy (`RPR`) is independent of the parental stock.
